### 2023-08-01 sl_report_pmcrc/paper0904.tex
updates

\documentclass[10pt]{article}
%\documentclass[twocolumn]{article}
%\documentclass[onecolumn]{article}
% \usepackage{scrtime} % for \thistime (this package MUST be listed first!)
\DeclareUnicodeCharacter{0301}{\'{e}}
\usepackage{times}
\usepackage{graphicx}
\usepackage{float}
\usepackage[margin=0.75in]{geometry}
\usepackage{fancyhdr}
\usepackage{caption}
\usepackage{notoccite}
\usepackage{pgfplotstable}
\usepackage{soul} %For highlights using \hl
%\usepackage[round]{natbib}
%\setcitestyle{aysep={}} %removes the comma between the author and year in citations
%\usepackage{underscore}
\usepackage{pdfpages}
\usepackage{xcolor,colortbl}%for changing cell colour
\usepackage[normalem]{ulem}
\useunder{\uline}{\ul}{}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{capt-of}
\pagestyle{fancy}
\setlength{\headheight}{15.2pt}
\setlength{\headsep}{13 pt}
\setlength{\parindent}{28 pt}
\setlength{\parskip}{12 pt}
\pagestyle{fancyplain}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
% \usepackage{color,amsmath,amssymb,amsthm,mathrsfs,amsfonts,dsfont}
\usepackage{xspace}
\usepackage{tikz-cd}
\usepackage{tikz}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{calc, arrows}
\usetikzlibrary{external}
\usepackage{pgfplots}
\pgfplotsset{layers/my layer set/.define layer set={background,main,foreground}{},
        set layers=my layer set,}

\usepackage{listings}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{tabulary}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

%for code
\lstdefinestyle{mystyle}{
	backgroundcolor=\color{backcolour},   
	commentstyle=\color{codegreen},
	keywordstyle=\color{magenta},
	numberstyle=\tiny\color{codegray},
	stringstyle=\color{codepurple},
	basicstyle=\ttfamily\footnotesize,
	breakatwhitespace=false,         
	breaklines=true,                 
	captionpos=b,                    
	keepspaces=true,                 
	numbers=left,                    
	numbersep=5pt,                  
	showspaces=false,                
	showstringspaces=false,
	showtabs=false,                  
	tabsize=2
}

\lstset{style=mystyle}
% \usetikzlibrary{pgfplots.clickable}
% \usepgfplotslibrary{clickable}
% tables
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{multirow}
% figs
\captionsetup[figure]{labelfont={color=blue}, font={color=black}}
%\usepackage{subfig}% http://ctan.org/pkg/subfig
\usepackage{subcaption}
%\newsubfloat{figure}% Allow sub-figures
\usepackage{caption}%lable fig caption as fig
%\captionsetup[subfigure]{labelfont=bf, justification=raggedright, labelformat=empty} %no caption label
\usepackage{stackengine} %places caption inside figure?
\captionsetup{subrefformat=empty} %when you reference the subcaption it will be (a) for example %{labelfont={color=blue}}
%\captionsetup[subfigure]{labelsep=colon}

% \usepackage{acronym}
% \usepackage{lineno}%for line numbers
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BIBLIOGRAPHY
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[backend=biber, giveninits=true, doi=false, isbn=false, natbib=true, url=true, eprint=false, style=authoryear-comp, sorting=nyt, sortcites=ynt, maxcitenames=2, maxbibnames=10, minbibnames = 10, uniquename=false, uniquelist=false, dashed=false]{biblatex} % can change the maxbibnames to cut long author lists to specified length followed by et al., currently set to 99.

%% bibliography for each chapter...
\DeclareFieldFormat[article,inbook,incollection,inproceedings,patent,thesis,unpublished]{title}{#1\isdot} % removes quotes around title
\renewbibmacro*{volume+number+eid}{%
	\printfield{volume}%
	%  \setunit*{\adddot}% DELETED
	\printfield{number}%
	\setunit{\space}%
	\printfield{eid}}
\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
%\renewcommand*{\newunitpunct}{\space} % remove period after date, but I like it. 
\renewbibmacro{in:}{\ifentrytype{article}{}{\printtext{\bibstring{in}\intitlepunct}}} % this remove the "In: Journal Name" from articles in the bibliography, which happens with the ynt 
\renewbibmacro*{note+pages}{%
	\printfield{note}%
	\setunit{,\space}% could add punctuation here for after volume
	\printfield{pages}%
	\newunit}    
\DefineBibliographyStrings{english}{% clears the pp from pages
	page = {\ifbibliography{}{\adddot}},
	pages = {\ifbibliography{}{\adddot}},
} 
\DeclareFieldFormat{journaltitle}{#1\isdot}
\renewcommand*{\revsdnamepunct}{}%remove comma between last name and first name
\DeclareNameAlias{sortname}{family-given}
% \DeclareNameAlias{sortname}{last-first}
\renewcommand*{\nameyeardelim}{\addspace} % remove comma in text between name and date
\addbibresource{ABC1.bib} % The filename of the bibliography
\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography
\renewrobustcmd*{\bibinitperiod}{}
% you'll have to play with the citation styles to resemble the standard in your field, or just leave them as is here. 
% or, if there is a bst file you like, just get rid of all this biblatex stuff and go back to bibtex. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% generally hyperref needs to be loaded last
\usepackage[hidelinks,colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
%\usepackage[hidelinks,colorlinks=false,citecolor=blue,urlcolor=darkbrown]{hyperref}
\tikzexternalize

\lhead{Synthetic Lethality} %This needs to change
\rhead{Alexander Turco}
\title{\sc Building sex-specific synthetic lethality network in cancer - Temp Name}
\author{\sc Alexander Turco}

\providecommand{\figref}[1]{Figure \ref{#1}}  %what?
\providecommand{\tabref}[1]{(Table \ref{#1})}  %what?
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}
\newcommand{\seg}{\texttt{Seg}\xspace}
\newcommand{\ecoli}{\mbox{\textit{E.\,coli}}\xspace}
\newcommand{\sclong}{\textit{Saccharomyces cerevisiae}\xspace}
\newcommand{\scshrt}{\mbox{\textit{S.\,cerevisiae}}\xspace}
\newcommand{\sce}{\mbox{\textit{S.\,cerevisiae}}\xspace}
\newcommand{\hslong}{\textit{Homo sapiens}\xspace}
\newcommand{\hsshrt}{\mbox{\textit{H.\,sapiens}}\xspace}
\newcommand{\hse}{\mbox{\textit{H.\,sapiens}}\xspace}
\newcommand{\celong}{\textit{Caenorhabditis elegans}\xspace}
\newcommand{\ceshrt}{\mbox{\textit{C.\,elegans}}\xspace}
\newcommand{\dmlong}{\textit{Drosophila melanogaster}\xspace}
\newcommand{\dmshrt}{\mbox{\textit{D.\,melanogaster}}\xspace}
\newcommand{\atlong}{\textit{Arabidopsis thaliana}\xspace}
\newcommand{\atshrt}{\mbox{\textit{A.\,thaliana}}\xspace}
\newcommand{\pflong}{\textit{Plasmodium falciparum}\xspace}
\newcommand{\pfshrt}{\mbox{\textit{P.\,falciparum}}\xspace}
%%%%BIBLIOGRAPHY

%Supplementary File Table Numbers:
\newcommand{\expdata}{S1\xspace}
\newcommand{\seqdata}{S2\xspace}
\newcommand{\protdata}{S3\xspace}
\newcommand{\blast}{S4\xspace}
\newcommand{\tabvar}{S5\xspace}
%Supplementary File Fig Numbers:

\newcommand{\expcor}{S1\xspace}
\newcommand{\expdistATCC}{S3\xspace}
\newcommand{\specialcell}[2][c]{%
	\begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
\newcommand{\beginsupplement}{%
	\setcounter{table}{0}
	\renewcommand{\thetable}{S\arabic{table}}%    %thetable references table counter 
	\setcounter{figure}{0}
	\renewcommand{\thefigure}{S\arabic{figure}}%
	\setcounter{equation}{0}
	\renewcommand{\theequation}{S\arabic{equation}}%

}
\renewcommand{\thesection}{}
\renewcommand{\thesubsection}{}
\renewcommand{\thesubsubsection}{}
\usepackage{setspace}
%adjust spacing
\doublespacing
\usepackage{titlesec}

\titlespacing\section{0pt}{12pt plus 2pt minus 2pt}{0pt plus 1pt minus 1pt}
\titlespacing\subsection{0pt}{12pt plus 2pt minus 2pt}{0pt plus 1pt minus 1pt}
\titlespacing\subsubsection{0pt}{12pt plus 2pt minus 2pt}{0pt plus 1pt minus 1pt}

% below 3 lines will put ALL table captions at the top...not sure if
% this is what we want but it is good enough for now

% \usepackage{float}
\floatstyle{plaintop}
\restylefloat{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

        \definecolor{atomictangerine}{rgb}{1.0, 0.6, 0.4}
        \definecolor{darkbrown}{rgb}{1.0, 0.56, 0.24}
        \colorlet{darkcol}{black!30!white}
        \colorlet{lightcol}{black!10!white}
        \definecolor{txtcol}{HTML}{F40000}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\widowpenalty10000
\clubpenalty10000

%\linenumbers %for line numbers
\onecolumn
%\twocolumn[  
%       \begin{@twocolumnfalse}
%               \begin{center}
                        \maketitle
%               \end{center}
%                       \bigskip

\thispagestyle{empty}
\noindent \textsuperscript{1} Department of, University, , ON, Canada

\newpage
\tableofcontents %Will add a nice Table of Contents
\newpage

\section{Abstract}

\newpage
\section{Introduction/Background Information}

	%What are they
	\subsection{What are Synthetic Lethal Interactions?}

	%Why are they important
	\subsection{Synthetic Lethal Interactions are Harnessed for Precision Oncology}

	%Building SL networks Pan-cancer
	\subsection{Building Pan-Cancer Synthetic Lethality Networks}

	%Bringing sex differences into the paper
	\subsection{Human Sex Differences add An Additional Layer of Complexity}

	%Bringing everything together
	\subsection{Building Pan-Cancer Synthetic Lethality Networks in a Sex Specific Manner}

\newpage
\section{Materials and Methods} 
%\label{methods}
%Custom scripts and commands utilized in this analysis can be found on \texttt{GitHub} at \url{https://github.com/opticrom/abcmcmc-thesis4c12}.

%\sloppy For a detailed protocol, see Supplementary files on \texttt{GitHub} at
 %\url{https://github.com/JohannaEnright/LCREntropyProject/}.

	\subsection{TCGA Data}

	RNA sequencing (RNA-seq) data was obtained from The Cancer Genome Atlas (TCGA). Raw STAR (Spliced Transcripts Alignment to a Reference) aligned counts for tumor tissue and healthy tissue samples were collected. The Cancer Genome Atlas contains genomic information which spans 33 cancer types. For the purpose of this study, only 12 of the 33 cancer types were considered. We first filtered out sex-biased cancers which include breast invasive carcinoma (BRCA), cervical cell carcinoma (CESC), ovarian serous cystadenocarcinoma (OV), prostate adenocarcinoma (PRAD), testicular germ cell tumors (TGCT), uterine corpus endometrial carcinoma (UCEC), and uterine carcinosarcoma (UCS). The reason for this was due to the fact that we are already aware of sex biases in these cancer types. Next, we filtered out blood cancers as well as cancers which lacked normal tissue gene expression samples. This included adrenocortical carcinoma (ACC), lymphoid neoplasm diffuse large b-cell lymphoma (DLBC), glioblastoma multiforme (GBM), acute myeloid leukemia (LAML), and brain lower grade glioma (LGG). The reason for this was due to the fact that we did not have adequate control samples to compare to tumor samples. Finally, we filtered out cancers that did not have any matching pairs of samples (NT and TP from same individual), as well as cancers that had less than 10 matched sample pairs across both males and females. This included mesothelioma (MESO), skin cutaneous melanoma (SKCM), thymoma (THYM), uveal melanoma (UVM), cholangiocarcinoma (CHOL), pancreatic adenocarcinoma (PAAD), pheochromocytoma and paraganglioma (PCPG), rectum adenocarcinoma (READ), and sarcoma (SARC). We selected matched tumor-normal sample pairs to help control for genetic background and other individual-specific factors that could influence gene expression. Once cancer types were selected, two pan-cancer raw count gene expression matrices were created, one for males and one for females.

	\begin{table}[ht]
	\centering
	\caption{List of 12 TCGA Cancer Types With Number of Matched Tumor-Normal Samples in Males and Females.}
	% Second version of table, with booktabs.
	\begin{tabulary}{\textwidth}{CCCC}\toprule
		& \multicolumn{3}{c}{TCGA information}
		\\\cmidrule(lr){2-4}
		& Cancer Type  & Matched Female Samples & Matched Male Samples\\\midrule
		& BLCA & 9 & 10 \\
		& COAD & 21 & 20 \\
		& ESCA & 5 & 8 \\
		& HNSC & 14 & 29 \\
		& KICH & 12 & 13 \\
		& KIRC & 20 & 52 \\
		& KIRP & 10 & 22 \\
		& LIHC & 22 & 28 \\
		& LUAD & 34 & 24 \\
		& LUSC & 14 & 37 \\
		& STAD & 10 & 23 \\
		& THCA & 42 & 17 \\
		& \textcolor{red}{TOTAL} & \textcolor{red}{213} & \textcolor{red}{283} \\\bottomrule
	\end{tabulary}
	\end{table}

	\subsection{Pre-filtering and Normalization of Raw RNA-seq Count Data}

	Raw count gene expression matrices for males and females were pre-filtered to remove genes unlikely to exhibit differential expression. For each matrix, we calculated the 90th quantile of overall gene expression as a threshold. For each gene in the matrix, we checked to see whether its expression was greater than the threshold in at least 1 sample. We removed genes where no samples showed an expression value greater than the quantile threshold.

	The pre-filtered matrices were then normalized using the DESeq2 package in R \citep{love2014moderated}. RNA-seq data must be normalized in order to account for factors that prevent the direct comparison of expression measures. The DESeq2 package employs a median of ratios normalization method to account for the inherent biases associated with RNA-seq data. Both sequencing depth (\# of reads generated per sample) and RNA composition (differences in composition of RNA molecules in a sample) are factors accounted for by the DESeq2 package. Raw counts are divided by size factors determined for each sample by computing the median ratio of gene counts relative to the geometric gene calculated per gene \citep{love2014moderated}. \figref{fig:1} summarizes the effects of normalization for both male and female RNA-seq data. The distribution of counts across samples becomes much more consistent, thus making the samples comparable.

	\begin{figure}[!h]
		\centering	
		\includegraphics[width=8.5cm, height=6cm]{normalied_male.png}\hfill
		\includegraphics[width=8.5cm, height=6cm]{normalied_female.png}
		\caption{Boxplots highlighting the distribution of raw RNA-seq data (top row) vs normalized RNA-seq data (bottom row) in males (left) and females (right) across normal tissue and tumor tissue samples from 12 TCGA cancer types. Each colour represents a specific cancer type.}
		\label{fig:1}
	\end{figure}

	\subsection{Data Quality Assessment (PCA, NPManova)}

	Once expression matrices were normalized, we performed principle component analysis (PCA) on the normalized counts to gain insights into potential factors contributing to the overall variance. Gene expression data is complex due to large number of variables (genes) present. PCA is a technique used to reduce dimensionality by transforming the data to a new set of variables (principle components) that summarize features of the data \citep{yeung2001principal}. In the context of this study, PCA is useful because it can take expression information from many genes, and reduce it down to fewer dimensions, making the data easier to explore. Using PCA, we explored the effects of tissue condition (normal vs tumor), and cancer type on gene expression.

	\begin{figure}[!h]
		\centering	
		\includegraphics[width=\textwidth,height=9cm]{all_cancerspca_analysis_condition_0.9.png}
		\caption{PCA plot highlighting the effect of tissue condition (normal vs tumor) on gene expression in males (left) and females (right) across 12 cancer types. Normal tissue samples are shown in red, tumor tissue samples are shown in blue.}
		\label{fig:2}
	\end{figure}

	\figref{fig:2} shows the effects of tissue condition (normal vs tumor) on gene expression across the 12 cancer types that were considered in this study. In both males and females, clusters based on tissue condition were present, however there was more variability in clusters of tumor tissue samples compared to clusters of normal tissue samples. \figref{fig:3} shows the effects of both tissue condition and cancer type on gene expression.

	\begin{figure}[!h]
		\centering	
		\includegraphics[width=\textwidth,height=9cm]{all_cancerspca_analysis_condition_cancertype_0.9.png}
		\caption{PCA plot highlighting the effect of tissue condition (normal vs tumor) and cancer type on gene expression in males (left) and females (right) across 12 cancer types.}
		\label{fig:3}
	\end{figure}

	When we added cancer type as a factor, we obtained much clearer clusters in the PCA plot, suggesting that cancer type contributes to the variation in gene expression.

	\subsection{Differential Gene Expression Analysis with DESeq2}

	\subsection{Fishers Exact Test For Synthetic Lethality}

\section{Results}

%	\subsection{Assessing the Evolution Simulator}

%	\subsection{Applying the Simulator to an ABC-MCMC}

\section{Discussion}

%	\subsection{Mutations Have the Ability to Destroy LCR Formation}

%	\subsection{Does the Length of a Repeat Play a Role in Insertions/Deletions}

%	\subsection{Why Parameters Could not be Estimated}

%	\subsection{Conclusion}

\clearpage\newpage
\section{Code \& Supplementary Materials}

The following files listed are located on the Graham Cluster at \texttt{$\sim$/projects/def-sushant/alextu/\newline slproject\_pmcrc}

Files used to obtain and format data can be found at \texttt{$\sim$/projects/def-sushant/alextu/\newline slproject\_pmcrc/raw\_data/scripts}
\begin{itemize}
\item\texttt{tcgabiolinks\_rawcount\_rnaseq\_download\_female.R}

This script utilizes the tcgabiolinks package in R to retrieve gene expression data (RNA-seq) for a specific set of TCGA barcodes. The TCGA barcodes used in this script came from females of a specific cancer type. For each cancer type, a raw gene expression matrix is produced with genes as rows, and samples as columns. DO ON LOGIN NODE, NOT ON COMPUTE NODE, COMPUTE NODE CANT DOWNLOAD

\item\texttt{tcgabiolinks\_rawcount\_rnaseq\_download\_female.R}

This script utilizes the tcgabiolinks package in R to retrieve gene expression data (RNA-seq) for a specific set of TCGA barcodes. The TCGA barcodes used in this script came from males of a specific cancer type. For each cancer type, a raw gene expression matrix is produced with genes as rows, and samples as columns.

\item\texttt{tcgabiolinks\_rawcount\_clinical\_download\_female.R}

This script utilizes the tcgabiolinks package in R to retrieve clinical metadata for a specific set of TCGA barcodes. The TCGA barcodes used in this script came from females of a specific cancer type. For each cancer type, a SummarizedExperiment (R object) is produced and clinical information such as age at diagnosis and tissue of origin can be extracted.

\item\texttt{tcgabiolinks\_rawcount\_clinical\_download\_male.R}

This script utilizes the tcgabiolinks package in R to retrieve clinical metadata for a specific set of TCGA barcodes. The TCGA barcodes used in this script came from males of a specific cancer type. For each cancer type, a SummarizedExperiment (R object) is produced and clinical information such as age at diagnosis and tissue of origin can be extracted.

\item\texttt{extract\_clinical\_data\_females.R}

This script extracts the clinical information of interest from the SummarizedExperiment (R object). We obtain barcode, tissue, and age information for females.

\item\texttt{extract\_clinical\_data\_males.R}

This script extracts the clinical information of interest from the SummarizedExperiment (R object). We obtain barcode, tissue, and age information for males. \newline

\item\texttt{extract\_both\_tp\_nt\_samples\_females.R}

This script

\item\texttt{extract\_both\_tp\_nt\_samples\_females.R}

This script

\item\texttt{merge\_rawcounts\_both\_tp\_nt\_samples\_females.R}

This script takes the gene expression matrices (with matched NT-TP samples) for females of each cancer type and merges them to create a raw pan-cancer matrix. Rows are genes and columns are samples. Samples from every cancer type utilized in the study are present in this matrix.

\item\texttt{merge\_rawcounts\_both\_tp\_nt\_samples\_males.R}

This script takes the gene expression matrices (with matched NT-TP samples) for males of each cancer type and merges them to create a raw pan-cancer matrix. Rows are genes and columns are samples. Samples from every cancer type utilized in the study are present in this matrix.

\end{itemize}

Files used to process and visualize data can be found at \texttt{$\sim$/projects/def-sushant/alextu/\newline slproject\_pmcrc/processed\_data/scripts}

\begin{itemize}
	\item\texttt{deseq2\_extract\_normalized\_counts\_allcancers\_female.R}

	This script

	\item\texttt{deseq2\_extract\_normalized\_counts\_allcancers\_male.R}

	This script

	\item\texttt{deseq2\_preprocessing\_boxplots\_quantile\_prefiltering\_allcancers\_female.R}

	This script

	\item\texttt{deseq2\_preprocessing\_boxplots\_quantile\_prefiltering\_allcancers\_male.R}

	This script

	\item\texttt{deseq2\_preprocessing\_boxplots\_quantile\_prefiltering\_female.R}

	This script

	\item\texttt{deseq2\_preprocessing\_boxplots\_quantile\_prefiltering\_male.R}

	This script

	\item\texttt{dge\_analysis\_deseq2.R}

	This script

	\item\texttt{gsea\_analysis\_postdeseq.R}

	This script

	\item\texttt{npmanova.R}

	This script

	\item\texttt{npmanova.R}

	This script

\end{itemize}

\clearpage\newpage
\section{References}

%%%FIGURES%%%%%

%%%%PRINTING BIBLIOGRAPHY%%%%
\nocite{*}
\printbibliography[heading=none, sorting=nyt]
\newpage

%\section{Appendix 1: Low Complexity Region Evolution Simulator}
%\texttt{/home/alext/scratch/abcmcmc-thesis4c12/mutations\_2\_BG\_vecs.cpp}
%\lstinputlisting[language=c++]{../mutations_2_BG_vecs.cpp}
%\newpage

%\section{Appendix 2: ABC-MCMC in C++}
%\texttt{/home/alext/scratch/abcmcmc-thesis4c12/abcmcmc.cpp}
%\lstinputlisting[language=c++]{../abcmcmc.cpp}
%\newpage

%\section{Appendix 3: Simulated Protein Summary Statistics}
%\texttt{/home/alext/scratch/abcmcmc-thesis4c12/abcmcmc.cpp/simulated\_protein.cpp}
%\lstinputlisting[language=c++]{../simulated_protein.cpp}
%\newpage

%\section{Appendix 4: Euclidean Distance and Normlization of Vectors}
%\texttt{/home/alext/scratch/abcmcmc-thesis4c12/abcmcmc.cpp/distance2.cpp}
%\lstinputlisting[language=c++]{../distance2.cpp}
%\newpage

%\section{Appendix 5: One Sample t-test}
%\texttt{/home/alext/scratch/abcmcmc-thesis4c12/abcmcmc.cpp/ttest.cpp}
%\lstinputlisting[language=c++]{../ttest.cpp}
%\newpage

\end{document}