### 2023-08-08 aug08meeting_pres4/may30ccgmeeting.tex
updates

% Inbuilt themes in beamer
\documentclass{beamer}

%Defining some colours
\definecolor{darkred}{rgb}{0.8,0,0}

% Theme choice: there a number of preset themes to choose from
% Play around with them, Cambridge is nice for first pres
%\usetheme{Szeged}
\usetheme{CambridgeUS} %setting the main theme
\usecolortheme{whale} % setting colour theme
\usefonttheme{professionalfonts} %font theme
\useinnertheme[shadow=true]{rounded}
%\useoutertheme{} %outer theme

%\setbeamertemplate{footline} %Remove footer line in all slides
\setbeamertemplate{navigation symbols}{} %removes navigation symbols
\setbeamertemplate{footline}[page number] %removes footer line, keeps pg#
\setbeamertemplate{caption}{\insertcaption}

%Setting colours for boxes and captions
\setbeamercolor{block title}{bg=blue!30, fg=black}
\setbeamercolor{block body}{bg=blue!10}
\setbeamercolor{frametitle}{fg=black}

%%%FOR VIDEOS
\usepackage{media9}
\usepackage{multimedia}
\usepackage{graphicx}
% For the Flowchart
%\usepackage{tikz}
%\usetikzlibrary{shapes.geometric, arrows}
%\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30]
%\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=orange!30]
%\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=green!30]
%\tikzstyle{arrow} = [thick,->,>=stealth]

%%% For the Timeline
\usepackage{xcolor}
\usepackage{tikz} \usetikzlibrary{calc, arrows.meta, intersections, patterns, positioning, shapes.misc, fadings, through,decorations.pathreplacing}

\definecolor{ColorOne}{rgb}{0.0,0.5,1.0} %Lightblue
\definecolor{ColorTwo}{rgb}{1.0,0.6,0.4} %lightorange
\definecolor{ColorThree}{rgb}{0.75,0.58,0.89} %lightpurple

% Title page details: 
\title[BEAP Dec 2022]{Double Trouble: Understanding Sex Differences in Synthetic Lethal interactions in Human Cancers \\~\\ Summer Wrap Up} 
\author{Alexander Turco}
\date{August 8, 2023}
\logo{\includegraphics[height=0.5cm, width=3cm]{logo.png}}

% Bibliography stuff
\usepackage[natbib=true, sorting=nyt, style=authoryear-comp]{biblatex}
\addbibresource{ABC1.bib}

%Extra packages
\usepackage{makecell}

%For itemize
\setbeamertemplate{itemize item}[triangle]

%For Flow Chart
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}

\tikzstyle{startstop} = [rectangle, 
minimum width=6cm, 
minimum height=1cm,
text centered, 
draw=black, 
fill=blue!30]

\tikzstyle{io} = [trapezium, 
trapezium stretches=true, % A later addition
trapezium left angle=70, 
trapezium right angle=110, 
minimum width=3cm, 
minimum height=1cm, text centered, 
draw=black, fill=blue!30]

\tikzstyle{process} = [rectangle, 
minimum width=3cm, 
minimum height=1cm, 
text align, 
text width=1cm, 
draw=black, 
fill=orange!30]

\tikzstyle{decision} = [diamond, 
minimum width=3cm, 
minimum height=1cm, 
text centered, 
draw=black, 
fill=green!30]
\tikzstyle{arrow} = [thick,->,>=stealth]

\begin{document}

	% For my introduction slides, there will be the slide with my title and name as well as an outline slide with the brief overview of what I will discuss.
	% Title page frame - SLIDE 1%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\section{Introduction}
	\begin{frame}
		\titlepage 
		\begin{center}
			%\includegraphics[width=7cm, height=4cm]{cellsatwar.png}
		\end{center}
	\end{frame}

	% Remove logo from the next slides
	\logo{}

	% Outline frame - SLIDE 2
	\begin{frame}{Overview}

		\begin{center}
		\begin{minipage}{6cm}

		  		\begin{block}{} Overall Objective and Approach \end{block}
		  		\begin{block}{} Synthetic Lethality Analysis Update \end{block}
		  		\begin{block}{} Next Steps/Future Work \end{block}
		  		\begin{block}{} Limitations of the study \end{block}
		  		\begin{block}{} Takeaways \& Tips for Future Undergrads \end{block}

		\end{minipage}
		\end{center}

	\end{frame}

	% For my Background slides, I will talk about important things such as WHat LCRs are, their evolution, stuff like that
	% SLIDE 3 - WHAT ARE LCRs%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	\section{Overall Objective and Approach}
	\begin{frame}{Overall Objective}
		\begin{itemize}
			%\includegraphics[width=11cm, height=6cm]{types_of_courses_taken.jpg}
			\item Can we build sex-specific synthetic lethality networks for various cancer types? \newline

			\item More specifically, we are trying to elucidate the differences in synthetic lethal interactions between males and females using a network based approach.
		\end{itemize}
	\end{frame}

	%%%%% APPRACH SECTION BEGINS HERE, 2 SLIDES

	\begin{frame}{Overall Project Workflow}
		\includegraphics[width=12cm, height=8cm]{project_workflow_slpmcrc_2.png}
	\end{frame}

	\section{Synthetic Lethality Analysis Update}
	\begin{frame}{Using DE Genes to Find Sex-Specific Synthetic Lethal Pairs}
		\begin{center}
			\includegraphics[width=7cm, height=4cm]{all_cancersvenndiagram_malede_vs_femalede_0.9.png}
		\end{center}
		\begin{enumerate}
			\item Find all potential gene pair combinations using differentially expressed genes \newline	
			\textcolor{blue}{${1289 \choose 2 } = 830,116$ Potential Male Pairs \newline}
			\textcolor{yellow}{${892 \choose 2 } = 397,386$ Potential Female Pairs}

			\item Identify pairs of genes that are BOTH perturbed (under and overexpressed) more frequently than INDIVIDUALLY perturbed using Fishers Exact Test
		\end{enumerate}
	\end{frame}

	\begin{frame}{Fishers Exact Test}
		Used to determine association between two categorical variables (overexpression and underexpression) when dealing with small sample sizes
		\begin{center}
		\includegraphics[width=10cm, height=7cm]{Contingency.png}
		\end{center}
	\end{frame}

	\begin{frame}{Ranking Genes by Number of Interactions}
		\includegraphics[width=12cm, height=5cm]{interactions.png}

		\begin{itemize}
		\item ISLE Method SL list Matches: Males = 14 pairs, Females = 11 pairs
		\item Experimentally Validated SL list Matches: Males = 21 pairs, Females = 2 pairs
		\end{itemize}
	\end{frame}

	\section{Next Steps/Future Work}
	\begin{frame}{Next Steps for Synthetic Lethality Analysis}
		\begin{enumerate}
			\item Refine the current list of synthetic lethal pairs by performing survival analysis to identify SL pairs that contribute to improved patient survival.
			\item Build male and female synthetic lethal interaction networks	
		\end{enumerate}
	\includegraphics[width=12cm, height=6cm]{network_interactions.png}
	\end{frame}

	\section{Limitations}
	\begin{frame}{A Major Limitation}
		In many cancers, normal tissues are not available in The Cancer Genome Atlas	
		\begin{center}
			\includegraphics[width=6cm, height=6cm]{lack.png}
		\end{center}
	\footnotetext[1]{\tiny Figure From  \cite{zeng2019selecting}}
	\end{frame}

	\section{Takeaways \& Tips}
	\begin{frame}{Experience Gained (Technical Skills)}

	\end{frame}

	\begin{frame}{Experience Gained (Personal Skills)}
		content...
	\end{frame}

	\begin{frame}{Major Takeaways for Future Undergrads}
		\begin{itemize}
			\item 
		\end{itemize}
	\end{frame}

	\begin{frame}{Thank You}
		To Sushant for taking a chance on me

		To everyone in the lab for the constant support and guidance

		11/10 would do again

		\begin{center}
			\includegraphics[width=7cm, height=5cm]{lab.jpg}
		\end{center}
	\end{frame}

\end{document}